Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})

### 2014-05-15 manuscript.tex
ongoing edits, problem with captions

\documentclass[12pt]{article}
\usepackage{geometry} 
\usepackage{hyperref}
\usepackage{color}
\usepackage[pdftex]{graphicx}  
%\usepackage{caption}
\usepackage{natbib}
\usepackage{authblk}
\usepackage{subscript}
\renewcommand{\baselinestretch}{1.5}
\bibliographystyle{apalike}

@@

%----------------------------------------
%AUTHORS
%----------------------------------------
\title{Natural variation in teosinte at the domestication locus teosinte branched1 (tb1)}

\author[1]{Laura Vann}
\author[1,2]{Thomas Kono}
\author[1,3]{Tanja Pyh\"aj\"arvi}
\author[1,4]{Matthew B. Hufford\thanks{mbhufford@iastate.edu}}
\author[1,5]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
\affil[1]{Department of Plant Sciences, University of California Davis}
\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
\affil[3]{Department of Biology, University of Oulu}
\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
\affil[5]{Center for Population Biology and Genome Center, University of California Davis}

@@

\jri{Check citation format for AJB. I don't think it's numbers. adjust bibliography style accordingly } \lev{I would rather send to an open access journal. I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for $500 which is cheaper than PLoS ONE}

@@

The lack of clustering of Hopscotch genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the Hopscotch insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Phyh\"aj\"arvi2013}. The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations. If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} \lev{Yes really 0.017 assignment to maize in the tb1 region, and avg assignment across chr1 is 0.2} \lev{ I also have a table of assignment values for SLO individuals based on genotype. Though we had decided this wasn't super informative because sample size was low} and much below the chromosome-wide average (0.20; \ref{Fig3Structure}).

@@

Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.  Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyh\"aj\"arvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings. \jri{deviations from HWE may be consistent too if we see excess of homozygotes. do we?}. This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyh\"aj\"arvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.  \jri{here we need a few sentences on selection.  the 5' UTR has much more negative D than the upstream. do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..

@@

% Figure S1. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS1LocusCartoon.pdf}
    \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.} 
\label{FigS1Locus}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Figure S2. Gel image
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS2Gel.png}
    \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
\label{FigS2Gel}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Figure S3. NJ Trees
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
    \caption{Neighbor-Joining tree of the sequenced region in the 5? UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}. Individuals with genotype data are colored: Homozygous for the teosinte (no Hopscotch) allele (red), homozygous for the maize (Hopscotch) allele (blue), heterozygotes (purple). TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the Hopscotch insertion. Black indicates individuals not genotyped for the Hopscotch insertion.} 
\label{FigS3NJ}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

@@

  \caption{Accessions of \emph{Zea mays} ssp. \emph{mexicana} (RIMME) and \emph{Zea mays} ssp. \emph{parviglumis} (RIMPA) sampled. RIHY is a \emph{Z. mays} ssp. \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mays} hybrid. }

@@

  \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}

@@

  \caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyh\"aj\"arvi2013}.}

@@

% Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
\begin{table}[htbp]
  \centering
  \caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
    \begin{tabular}{ccccc}
    \toprule
    \textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
    \midrule
    Ejutla A & 4     & 0.15217 & 0.11902 & 0.76191 \\
    Ejutla B & 5     & 0.15258 & 0.14877 & 0.07412 \\
    La Mesa & 3     & 0.12802 & 0.08926 & 1.09209 \\
    San Lorenzo & 3     & 0.09098 & 0.08926 & 0.04845 \\
    \bottomrule
    \end{tabular}%
  \label{TableS4DiversitySNP}%
\end{table}%

>>>>

\documentclass[12pt]{article}
\usepackage{geometry} 
\usepackage{hyperref}
\usepackage{color}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[pdftex]{graphicx}  
%\usepackage{caption}
\usepackage{natbib}
\usepackage{authblk}
\usepackage{subscript}
\renewcommand{\baselinestretch}{1.5}
\bibliographystyle{apalike}

@@

%----------------------------------------
%AUTHORS
%----------------------------------------
\title{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}

\author[1]{Laura Vann}
\author[1,2]{Thomas Kono}
\author[1,3]{Tanja Pyh\"aj\"arvi}
\author[1,4]{Matthew B. Hufford\thanks{mhufford@iastate.edu}}
\author[1,5]{Jeffrey Ross-Ibarra\thanks{rossibarra@ucdavis.edu}}
\affil[1]{Department of Plant Sciences, University of California Davis}
\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota Twin Cities}
\affil[3]{Department of Biology, University of Oulu}
\affil[4]{Department of Ecology, Evolution, and Organismal Biology, Iowa State University}
\affil[5]{Center for Population Biology and Genome Center, University of California Davis}

@@

\jri{Check citation format for AJB. I don't think it's numbers. adjust bibliography style accordingly } \lev{I would rather send to an open access journal. I don't like that no everyone would be able to have access to paper} \jri{AJB allows open access for \$500 which is cheaper than PLoS ONE}

@@

The lack of clustering of Hopscotch genotypes in our NJ tree as well as the lack of LD around \emph{tb1} does not support the hypothesis that the Hopscotch insertion in these populations of \emph{parviglumis} is the result of recent introgression. However, to further explore this hypothesis we performed a STRUCTURE analysis using Illumina MaizeSNP50 data from four of our \emph{parviglumis} populations (EjuA, EjuB, MSA, and SLO) and the maize 282 diversity panel \citep{Cook2012, Pyhajarvi2013}. The linkage model implemented in STRUCTURE can be used to identify ancestry of blocks of linked variants, which would arise as a result of recent admixture between populations. If the Hopscotch insertion is present in populations of \emph{parviglumis} as a result of recent admixture with domesticated maize, we would expect the insertion and linked variants in surrounding sites to be assigned to the "maize" cluster in our STRUCTURE runs, not the "teosinte" cluster. In all runs, assignment to maize in the \emph{tb1} region across all four \emph{parviglumis} populations is low (average 0.017) \jri{is this really 0.017 or 0.17?} \lev{Yes really 0.017 assignment to maize in the tb1 region, and avg assignment across chr1 is 0.2} \lev{ I also have a table of assignment values for SLO individuals based on genotype. Though we had decided this wasn't super informative because sample size was low} and much below the chromosome-wide average (0.20; \ref{Fig3Structure}).

@@

Other explanation for differential frequencies of the Hopscotch among teosinte populations include both drift and natural selection.  Previous studies using both SSRs and genome-wide SNP data have found evidence for a population bottleneck in the San Lorenzo population \citep{Hufford2010, Pyhajarvi2013}, and the lower levels of sequence diversity in the 5' UTR region and the more positive values of Tajima's D we present here are consistent with these findings. \jri{deviations from HWE may be consistent too if we see excess of homozygotes. do we?}. This bottleneck, however, does not explain differences in Hopscotch allele frequency among populations, and the available information on diversity and population structure among these populations \citep{Hufford2010, Pyhajarvi2013} is not suggestive of colonization or other demographic events that might predict a high frequency of the allele in multiple populations.  \jri{here we need a few sentences on selection.  the 5' UTR has much more negative D than the upstream. do we know the Hop genotype for sequenced lines? can we separate the sequences into hop/no hop and look for differences? it wasn't until we did this that gt1 stuff really popped out.} \lev{we should know for some of them, i will check}
%   from expected genotype frequencies under Hardy-Weinberg equilibrium at the Hopscotch insertion in the San Lorenzo population and additional populations may be indicative of demographic changes that have affected genome-wide patterns of variation, or may be the result of selective pressures acting on or near the \emph{tb1} locus..

@@

% Figure S1. Cartoon representation of tb1 locus
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS1LocusCartoon.pdf}
   % \caption{Representation of the upstream regulatory region of \emph{tb1}, showing the \emph{tb1} coding region (green) and the Hopscotch insertion (red). Arrows show the location of primer sets; in black, primers used for amplification and sequencing (Sequenced region 1; within the 5' UTR, and sequenced region 2; 66,169 bp upstream from the tb1 ORF); in blue, primers used to genotype the Hopscotch insertion.}
\label{FigS1Locus}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Figure S2. Gel image
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS2Gel.png}
   % \caption{Agarose gel image of amplification products using our primer sets. Genotypes are indicated at the top of the gel.} 
\label{FigS2Gel}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

% Figure S3. NJ Trees
%-------------------------------------------------------------------
\begin{figure*}[!t]
  \begin{center}
   \includegraphics[width=150mm]{FigS3NJtrees.pdf}
    %\caption{Neighbor-Joining tree of the sequenced region in the 5? UTR (right; sequenced region 1) and the 66,169 bp upstream region (left; sequenced region 2) of \emph{tb1}. Individuals with genotype data are colored: Homozygous for the teosinte (no Hopscotch) allele (red), homozygous for the maize (Hopscotch) allele (blue), heterozygotes (purple). TILs (teosinte inbred lines) are colored in green, with stars indicating the 3 TILs known to have the Hopscotch insertion. Black indicates individuals not genotyped for the Hopscotch insertion.} 
\label{FigS3NJ}
  \end{center}
\end{figure*}
%-------------------------------------------------------------------

@@

  %\caption{Accessions of \emph{Zea mays} ssp. \emph{mexicana} (RIMME) and \emph{Zea mays} ssp. \emph{parviglumis} (RIMPA) sampled. RIHY is a \emph{Z. mays} ssp. \emph{parviglumis} and \emph{Zea mays} ssp. \emph{mays} hybrid. }

@@

 % \caption{\emph{Zea mays} ssp. \emph{mays} (RIMMA) sampled for genotyping}

@@

  %\caption{Variables and rotations used for the 6 principal components used for BayEnv calculations and their corresponding Bayes Factors. Modified from \citet{Pyhajarvi2013}.}

@@

% Table S4 Diversity stats in tb1 region based on SNP 50 genotyping data
\begin{table}[htbp]
  \centering
  %\caption{Diversity in the \emph{tb1} region based on the maize SNP50 genotyping data}
    \begin{tabular}{ccccc}
    \toprule
    %\textbf{Population} & \textbf{\# seg sites} & \textbf{$\theta_{\subscript{\pi}}$ / bp} & \textbf{$\theta_{\subscript{W}}$ / bp} & \textbf{Tajima?s D} \\
    \midrule
    Ejutla A & 4     & 0.15217 & 0.11902 & 0.76191 \\
    Ejutla B & 5     & 0.15258 & 0.14877 & 0.07412 \\
    La Mesa & 3     & 0.12802 & 0.08926 & 1.09209 \\
    San Lorenzo & 3     & 0.09098 & 0.08926 & 0.04845 \\
    \bottomrule
    \end{tabular}%
  \label{TableS4DiversitySNP}%
\end{table}%
